---
title: "ECo 602 - Analysis of Environmental Data"
subtitle: "Final Project Template: Data Analysis"
author: "Sarah Traore"
date: "Fall 2020"
output:
  html_document:
    theme: readable
    toc: TRUE
    toc_float: TRUE
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
options(knitr.duplicate.label = TRUE)
```


Everything above this line is the header.

<!-- The following text won't be displayed in your document.  It tells R how to make nicer looking buttons for your tabbed content. -->

<style type="text/css" rel="stylesheet">

.btn {
    border-width: 0 0 0 0;
    font-weight: normal;
    text-transform: none;
}

.btn-default {
    color: #2ecc71;
    background-color: #ffffff;
    border-color: #ffffff;
}
</style>


Use this document as a template to build your final project.

If you've already created the file index.Rmd in your docs folder, you can paste everything below the header into that document.

If you don't yet have an index.Rmd file, save this document as index.Rmd in the docs folder of your main class folder.

Edit this template to fill in your own content.


# Introduction

Type a brief introduction here.


# Functions 1: Data Structure Functions {.tabset .tabset-pills}

Look at the source code for the final project template, you'll notice the text:

{.tabset .tabset-pills}

This creates tabbed content for all the following sections that are one-level below.

- If you include them on a primary header (one #), all of the subsequent secondary headers will in tabs.




## `c()`

The following is the markdown text needed to reproduce my code example for `c()`:


The function `c()` *combines* or *concatenates* its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

- All of the elements must be of the same *type*.
  - I can't combine character and numeric types in the same call to `c()`

Here's two examples using numeric and character data types:

```{r}
# Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

# Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```

I can show the contents of a vector by typing the name of the vector, or using the `print()` function.

```{r}
# Typing the name of the vector into the console prints the contents
num_vec

# The print() function accomplishes the same task:
print(char_vec)
```


## `data.frame()`

Type your data frame example here.
```{r data_frame_example_1}
# build a data frame from scratch
dat_1 = data.frame(
  letters = c("b", "w", "B"),
  numbers_1 = c(34, 6, 123454))
# print the content of the data frame
dat_1
# subset a column by name
dat_1$letters
```


## `matrix()`

Type your matrix example here.


## `length()`


## `nrow()`


## `ncol()`


## `dim()`


# Functions 2: Numerical Data Exploration  {.tabset .tabset-pills}

## `summary()`

## `mean()`

## `sd()`
 


# Functions 3: Graphical Data Exploration {.tabset .tabset-pills}

See the final project description for a list of the required functions.


```{r scatterplot_1, fig.asp=0.75}
require(palmerpenguins)

plot(bill_length_mm ~ body_mass_g, data = penguins, pch = 16)

```

# Functions 4: Distribution Functions {.tabset .tabset-pills}

See the final project description for a list of the required functions.



# Functions 5: Other Functions {.tabset .tabset-pills}

See the final project description for a list of the required functions.




# Part 2: Data Anaylsis
#Data Structure
```{r}
require(here)
dat_delomys = read.csv(here("data", "delomys.csv"))


```

# Data Exploration
# Numerical Exploration
```{r Data Exploration}

require(here)
dat_delomys = read.csv(here("data", "delomys.csv"))

summary(dat_delomys$body_mass)

summary(dat_delomys$body_length)
#Mean
mean(dat_delomys$body_mass)
mean(dat_delomys$body_length)

#Standard deviation
sd(dat_delomys$body_mass)
sd(dat_delomys$body_length)

# Test of normality
shapiro.test(dat_delomys$body_mass)
shapiro.test(dat_delomys$body_length)

```

# Graphical Exploration

```{r Graphical Exploration}

#Scatterplot

plot(body_length ~ body_mass, data = dat_delomys,
     main = "Scatterplot of body \n length & body mass", col='black',
     pch = 1, cex = 1, xlab = "Body mass", ylab = "Body length",
     xlim = c(0, 110), ylim =c(0,266))
```
```{r}
#Histograms of body mass & body length
par(mfrow = c(2,2))
hist(dat_delomys$body_mass, breaks = 40)


hist(dat_delomys$body_length, breaks = 40)
```


```{r Graphical Exploration}
#Conditonal boxplots

par(mfrow = c(2,2))
boxplot(body_mass ~ sex, data = dat_delomys, 
        main = "Conditional boxplot of \n body mass by sex")
boxplot(body_mass ~ binomial, data = dat_delomys, 
        main = "Conditional boxplot of \n body mass grouped by species")
boxplot(body_mass ~ binomial * sex, data = dat_delomys,
        main = "Conditional boxplot of \n body mass grouped by species & sex")
```
# Model Building
```{r}
#Model 1: Simple linear regression body_length ~ body_mass
fit_1 = lm(body_length ~ body_mass, data = dat_delomys)
summary(fit_1)

# Model 2: 1-way ANOVA body_mass ~ sex
fit_2 = lm(body_mass ~ sex, data = dat_delomys)
summary(fit_2)

# Model 3: 1-way ANOVA body_mass ~ binomial
fit_3 = lm(body_mass ~ binomial, data = dat_delomys)
summary(fit_3)

#Model 4: 2-way additional ANOVA body_mass ~ sex + binomial
fit_4 = lm(body_mass ~ sex + binomial, data = dat_delomys)
summary(fit_4)

# Model 5: 2-way ANOVA factorialbody_mass ~ sex *binomial
fit_5 = lm(body_mass ~ sex * binomial, data = dat_delomys)
summary(fit_5)

```
# Model Diagnostics

```{r}
#Histograms of the residuals 

hist(residuals(fit_1), breaks = 40)

hist(residuals(fit_2), breaks = 40)

hist(residuals(fit_3), breaks = 40)

hist(residuals(fit_4), breaks = 40)

hist(residuals(fit_5), breaks = 40)
```

The null hypothesis is:" the residuals are drawn from a normally-distributed population"
Let's test the null hypothesis for each model

```{r}
#Model 1 
shapiro.test(residuals(fit_1))

#Model 2 
shapiro.test(residuals(fit_2))

#Model 3 
shapiro.test(residuals(fit_3))

#Model 4 
shapiro.test(residuals(fit_4))

#Model 5 
shapiro.test(residuals(fit_5))
```
Answer question

# Model Interpretation
# Body Length
```{r  Model 1 coefficient table}


knitr::kable(coef(summary(fit_1)))

```

Answer questions

# Body Mass: Model Coefficients
```{r}
#Model 2 
knitr::kable(coef(summary(fit_2)))

#Model 3 
knitr::kable(coef(summary(fit_3)))

#Model 4 
knitr::kable(coef(summary(fit_4)))

#Model 5 
knitr::kable(coef(summary(fit_5)))

```


# Body Mass: ANOVA
```{r ANOVA table}

#Model 2 
knitr::kable(anova(fit_2))

#Model 3 
knitr::kable(anova(fit_3))

#Model 4
knitr::kable(anova(fit_4))

#Model 5 
knitr::kable(anova(fit_5))
```
Answer questions

# Model Comparison: Body Mass

We use the AIC criterion to compare the models.

```{r}
AIC_2 = AIC(fit_2)

AIC_3 = AIC(fit_3)

AIC_4 = AIC(fit_4)

AIC_5 = AIC(fit_5)
```
Model 4 and 5 the 2-way ANOVA have the lowest AIC 
I would choose the 2-way factorial ANOVA because

