---
title: "In class example"
author: "Sarah Traore"
date: "10/21/2020"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Making a new code chunk

ctrl+alt+i is a shortcut to create a chunk

## Reading a csv file

I'm going to read some data from a csv (comma-separated-values) file.

I can use read.csv() along with the path to the file to read the data into a data.frame.

Here's an example:
  
  
  ```{r read_birds_data}

require(here)

dat_birds = read.csv(here("data", "hab.sta.csv"))

```



## Previewing the contents of a data.frame



I can use the head() function to print the first few lines of a data.frame.



Here's an example with the birds data



```{r}

head(dat_birds)


```

## Demo plot

I can use plot to create a histogram
```{r simple plot}
hist(dat_birds$aspect)

```


## How to customize a histogram

```{r custom histogram 1}
hist(dat_birds$aspect, main = "Histogram of study site aspect", breaks = 36)
```

